Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-07 doc/report/msc-report.tex
added a pair of TODOs and a future work idea.
doc/report/msc-report.tex: * corrections of sentences and misspellings. * added to and polished Aggregates section. * changed "defined" to "written" for a more clear discussion of mappers.

This thesis, fully, describes the Zohmg system and how to use it.

Chapter 2 the theoretical base of the problem statement is established.

@@

Chapters 7 and 8 discusses the results, related, and future work, and
presents the conclusion, respectively.

@@

The school book example of a data cube is three dimensional where each of
the three dimensions correspond to one of the attributes time, products,
and stores. Adding another attribute adds another dimension to the data
cube, because of the one to one relationship of attributes and dimensions.

@@

An aggregate is a composite of several values. Typical examples of aggregate functions are average, count and sum.

In a standard OLAP setup, a number of aggregates are pre-computed and stored in the cube. These base aggregates represent only a fraction of the very many possible aggregations. The remaining aggregates are computed from the base aggregates as they are demanded.

@@

Depending on how the rowkeys are formatted different hierarchies are
created. The order in which the data is stored is very important, it
defines how effective scanning for data will be. For instance, let's say
we have thousand rows and want ten out of those. In this case only one per
cent of the data is interesting to us, if a scan would be required to
visit all the thousand rows a lot of rows are skipped. Skipping rows is
expensive in that sense that we have to read them but throw away the
result. The goal is to push this cost to a bare minimum, using as many as
possible of the rows we scan.

To minimize the cost of scanning for data we want it to be ordered so that
a scan would use every row it visits. In the case with thousand rows of
which we want ten, it would mean that we just scan ten rows and use all of
them.

@@

Both Hadoop and HBase are written in Java---a popular high-level
object-oriented programming language.

@@

Python is a popular general-purpose high-level multi-paradigm---most
notably: object-oriented, imperative and functional---scripting language.
\cite{python}

@@

A simple, fast, complete and extensible JSON encoder and decoder for
Python. The Python module simplejson is implemented entirely in Python and
has no external dependencies. However a C extension which provides a
serious speed boost is included. \cite{simplejson}

@@

@@

The user's data sets are defined using YAML. The data set files reside in
the config directory.

@@

The user defines a mapper which Zohmg wraps, interpreting the output from the
mapper program and storing it in HBase.

@@

The data export server is uilt with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URI it was fed.

@@

\noindent Edit the configuration files.

@@

Import jobs use mappers defined by the user, either in Python or Java.
Mappers written by the user import each line from the input file and set
the arguments to the map function, key and value, as line number and line
contents, respectively. The user can then perform computations on these
entities and finally yield the results as a dictionary. The resulting
dictionary will be interpreted by Zohmg and stored accordingly.

@@

The data server have three methods of serving data: raw, transformed, and
static. The raw and transformed data are aggregates served from HBase,
transformed in the latter case, while the static method just serve static
files from the project's \textit{clients} directory. Aggregates served from
HBase are returned to the user as JSONP.

@@

It is possible to build static web pages out of HTML and JavaScript, which
can interact with the data server. These static pages can for instance
present some sort of query building mechanism (for insntance with drop down
boxes) and then send this query to the data server and graph the resulting
data.

>>>>

This thesis fully describes the Zohmg system and how to use it.

Chapter 2 establish the theoretical base of the problem statement.

@@

Chapters 7 and 8 discuss the results, related, and future work, and
presents the conclusion, respectively.

@@

The school book example of a data cube is three dimensional where each of
the three dimensions correspond to one of the attributes time, products,
and stores. Adding another attribute, for instance sales, adds another
dimension to the data cube, because of the one to one relationship of
attributes and dimensions.

@@

An aggregate is a composite of several values. Typical examples of
aggregate functions are average, count, and sum.

In a standard OLAP setup, a number of aggregates are pre-computed and
stored in the cube. These base aggregates represent only a fraction of the
many possible aggregations. The remaining aggregates are computed from the
base aggregates as they are demanded.

The main reason for pre-computing aggregates is to reduce access times, the
time it takes to compute an aggregate may be unacceptable to a waiting
user. Pre-computing frequently requested aggregates is therefore regularly
done in the background, giving the user fast access to these.
\cite{olap_solutions}

The classic trade-off between time and space applys to wether or not
pre-compute an aggregate. Pre-computing to many aggregates might render an
unfeasible of data amount to store, while pre-computing to few aggregates
renders longer access times for aggregates which are not pre-computed. This
problem is further discussed in the thesis, but left to the user of the
Zohmg system to solve.

@@

Depending on how the rowkeys are formatted different hierarchies are
created. The order in which the data is stored is important, it defines how
effective scanning data will be. In the case of having thousand rows and
requesting ten out of these, then only one por cent of the data is
interesting. If a scan would be required to visit all the thousand rows a
lot of rows are skipped. Skipping rows is expensive in that sense that they
are but the result is thrown away. The goal is to push this cost to a bare
minimum, using as many of the rows as possible.

The minimum cost of scanning data is achieved by ordering the data so that
a scan uses every visited row. In the case with thousand rows of which ten
are wanted, would mean that ten rows scan are scanned of which all are
used.

@@

Both Hadoop and HBase are written in Java---a high-level object-oriented
programming language.

@@

Python is a general-purpose high-level multi-paradigm---amongst others:
object-oriented, imperative and functional---scripting language.
\cite{python}

@@

A simple, fast, complete and extensible JSON encoder and decoder for
Python. The Python module simplejson is implemented entirely in Python and
has no external dependencies. However a C extension which provides a
serious speed boost is included in simplejson. \cite{simplejson}

@@

Using Zohmg the user first creates a project directory. This is further
explained in chapter 5.

@@

The user's data sets are defined using YAML. The data set files reside in
the project's \texttt{config} directory.

@@

The user writes a mapper which Zohmg wraps, interpreting the output from
the mapper program and storing it in HBase.

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URI it was fed.

@@

\noindent Edit the configuration files, which are text files in YAML and
Python format.

@@

Import jobs use mappers written by the user, either in Python or Java.
Mappers written by the user import each line from the input file and set
the arguments to the map function, key and value, as line number and line
contents, respectively. The user can then perform computations on these
entities and finally yield the results as a dictionary. The resulting
dictionary will be interpreted by Zohmg and stored accordingly.

@@

The data server have three methods of serving data: raw, transformed, or
static. The raw and transformed data are aggregates served from HBase,
transformed in the latter case, while the static method just serve static
files from the project's \texttt{clients} directory. Aggregates served from
HBase are returned to the user as JSONP.

@@

It is possible to build static web pages out of HTML and JavaScript, which
can interact with the data server. These static pages can for instance
present some sort of query building mechanism---for instance with drop down
boxes---and then send this query to the data server and graph the resulting
data.
